\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}

I would like to thank Marek for giving me the opportunity to conduct my research in the IceCube group at DESY. It was a pleasure to work in this group that combines the various research topics performed in IceCube but also offers the most friendly and welcoming atmosphere and wonderful colleagues.

I especially want to thank Summer for her continuous support and guidance throughout my five years at DESY. Thank you for all the invaluable and honest feedback on my work and beyond. I am grateful for all the opportunities you gave me and for everything I learned from you. I am very happy that I had the chance to work with you!

I would also like to thank Alex who shared 2L01 with me throughout most of my time at DESY and was a precious resource of insights, new ideas, and prompt support on anything I was struggling with.

Special thanks to Carlos, who was part of initiating this project and who was always there to help me with any questions I had. Thanks for inviting me over to the US and for the great time we had there.

Thanks to Cris, JVS, Alex, Neha, Richard, Navid, and Summer for proofreading my thesis in order to make it a nice read.

Thanks to the desy family for making my time here so enjoyable, especially the ex-2L04 gang, you're the best!

Thanks to my family and friends for their continuous support and encouragement. 

Finally, thank you Charlotte for always being there for me and for all the love and support you give me. I could not have done this without you!


\begin{flushright}
	\textit{Leander Fischer}
\end{flushright}