\thispagestyle{plain}
\begin{center}
	\vspace*{1cm}

	\LARGE
	\textbf{First Search for Heavy Neutral Leptons with IceCube DeepCore}
	\large

	\vspace{0.8cm}

	\textbf{Dissertation} \\
	zur Erlangung des akademischen Grades \\
	doctor rerum naturalium \\
	(Dr. rer. nat.) \\

	\vspace{0.5cm}

	im Fach: Physik \\
	Spezialisierung: Experimentalphysik \\

	\vspace{0.5cm}

	eingereicht an der \\
	Mathematisch-Naturwissenschaftlichen Fakultät \\
	der Humboldt-Universität zu Berlin \\

	\vspace{0.5cm}

	von \\
	\textbf{Leander Fischer M. Sc.} \\

	% uncomment for publication
	\vspace{0.8cm}

	% % commented out for publication
	% geboren am 24. Oktober 1992 \\
	% in Heidelberg

	\vspace{0.5cm}

	Präsidentin der Humboldt-Universität zu Berlin \\
	Prof. Dr. Julia von Blumenthal \\

	\vspace{0.5cm}

	Dekanin der Mathematisch-Naturwissenschaftlichen Fakultät \\
	Prof. Dr. Caren Tischendorf \\

	\vspace{0.5cm}

	% uncomment for publication
	\vspace{0.8cm}
	Gutachter:innen: \\
	Prof. Dr. Marek Kowalski, Humboldt-Universität zu Berlin \\
	Prof. Dr. Joachim Kopp, Johannes Gutenberg-Universität Mainz \\
	Dr. Priscilla Pani, DESY Zeuthen \\

	\vspace{0.5cm}

	Tag der mündlichen Prüfung: \\
	08.07.2024

\end{center}

\newpage
\thispagestyle{plain}

\begin{flushleft}	
	\vspace*{8cm}
	
	\textbf{Copyright Notice} \\
	This work is licensed under CC BY-SA 4.0.

	To view a copy of this license, visit: \\\url{https://creativecommons.org/licenses/by-sa/4.0/}

	\medskip
	
	\textbf{Colophon} \\
	This document was typeset with the help of \href{https://sourceforge.net/projects/koma-script/}{\KOMAScript} and \href{https://www.latex-project.org/}{\LaTeX} using the open-source \href{https://github.com/fmarotta/kaobook/}{kaobook} template class. \\
	
	\medskip
	
	The source code of this thesis is available  at: \\\url{https://github.com/LeanderFischer/phd_thesis}

	\medskip

\end{flushleft}
